---
title: "Empirical Project 1"
subtitle: "Tillman Degens"
author: "Coauthored by Joey and Devin"
date: "4/29/2021"
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
output: 
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: true
fig_caption: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

>
It is extremely rare to be able to conduct a randomized experiment on the division of medical insurance and its impact on health care use and outcome. But, in March 2008, the state of Oregon did just that. In the Oregon Health Plan Experiment, 6387 individuals were selected to receive Medicaid in a unique lottery. Researchers were able to conduct interviews 25 months after the lottery occurred to determine its impact on lottery winner’s health and usage of health care. This project aims to conduct an exploratory analysis on a potential selection bias for the treatment and control group, as well as provide a basic summary of the impact of the program on important health outcomes. The results indicate that for the most part participants were randomly assigned to the treatment and control group. They also indicate that the policy had a larger impact in increasing health care usage (such as doctors’ visits, or prescriptions taken) compared to health outcomes. This study was significant due to the policy debate around public health care and its impact on the economy. 

\pagebreak 

# Questions to answer

## Question 1
>Explain the difference between the variables treatment and ohp_all_ever_survey. Explain why treatment is the treatment variable (Di), rather than ohp_all_ever_survey.

The difference between the  treatment and ohp_all_ever_survey is that treatment is individuals that 
were entered into medicaid specifically through the lottery, whereas ohp_all_ever_survey are those that
have been enrolled in medicaid independent of the lottery.

```{r}
#Load packages and libraries
library(pacman)
library(kableExtra)
p_load(readr,dplyr, tidyverse,
       ggplot2, skimr, haven, stargazer,
       tidymodels, skimr, janitor, magrittr,
       datasets, rpart.plot, baguette, glmnet,
       tune, haven, ranger, data.table, parallel,
       sandwich, modelsummary,  huxtable, hrbrthemes)
```


## Question 2
>Provide evidence that the OHP lottery really did randomly assign individuals to treatment
and control groups. Similar to Table 1 in Taubman et al. (2014), please create a nicely
formatted table that reports means of 4 to 6 relevant characteristics for individuals in the
control group.
Note: Part of this question is to get you to think about which variables should be balanced
in a randomized experiment. You need to read carefully through all the variables in the
dataset (documentation attached at the end of this file) and decide which 4 to 6 you will
summarize.

I selected Gender, Education, Race (non-white), Number of Doctors Visits, Medicine, and Cholesterol. I selected these variables because I believe that they all explain how an individual will interact with the health care system and their health outcome. Age was selected because it impacts the price of private health insurance as well as probability of an underlying condition. Gender was selected as it changes the distribution of health conditions (i.e. breast cancer). Race and education were chosen as societal factors and systematic discrimination have been proven to influence how individuals use health care. 

## Question 3
>For each of the variables you summarized above, calculate: (i) the difference between the mean in the treatment group and the mean in the control group. (ii) the standard error for the difference in means.
Add these as columns two and three to the table you started in question 2.

```{r}
setwd("C:/Users/tillm/OneDrive/Desktop/525")
#load data from downloads since Rconsole cannot download ".dta" files
ohp_data <- read_dta("ohp.dta")
```

```{r, results='asis', message=FALSE}
Gender = lm(gender_inp ~ treatment, data = ohp_data)
Age = lm(age_inp ~ treatment, data = ohp_data)
Not_White = lm(race_nwother_inp ~ treatment, data = ohp_data)
Education = lm(edu_inp ~ treatment, data = ohp_data)
Medicine= lm(rx_num_mod_inp ~ treatment, data = ohp_data)
Cholesterol = lm(chl_inp ~ treatment, data = ohp_data)

stargazer(Gender, Age, Not_White,Education, Cholesterol,
          title = "Overall Characterisitics for Explanatory Variables",
          font.size = "small",
          omit.stat = c("f", "ser"),
          column.sep.width = "6pt")


```

The Constant row in this table represents the mean of the control group. The treatment row represents the difference in means between the control group and the treatment group. The numbers in parenthesis below the treatment row represent the standard errors of the differences in means.The Difference in Group Means  = Treatment Effect +Selection Bias. Here we are looking to see if the explanatory variables that would influence health outcomes are properly randomized. If that is the case we can assume that selection bias is nonexistent.

```{r, results='asis', echo= FALSE}
ohp_sum_2 <- ohp_data %>%
    group_by(treatment) %>%
    summarise(
      Gender = mean(gender_inp),
      Age = mean(age_inp, na.rm = TRUE),
      Not_White = mean(race_nwother_inp, na.rm = TRUE),
      Education = mean(edu_inp, na.rm= TRUE),
      Cholesterol = mean(chl_inp, na.rm = TRUE))

#Turn the summary table into a dataframe
ohp_sum_df = as.data.frame(ohp_sum_2)

#Pivot longer for better readability
ohp_sum_df = ohp_sum_df %>% pivot_longer(!treatment)

#Filter treatment and control to begin to reshape table
treatment1 = ohp_sum_df %>% filter(treatment==1)
treatment0 = ohp_sum_df %>% filter(treatment == 0 )

#Join to get two rows
working_table = treatment1 %>% left_join(treatment0, by = "name")

#Select relevant rows
working_table = working_table %>% select(name, value.x, value.y)


#Create a variable for "difference in means"
working_table = working_table %>% mutate(difference = value.x-value.y)
# Rename variables 
working_table = working_table %>% rename("Characteristic"= name,
                                         "Control"= value.y, "Treatment" = value.x)


#Filter out treatment and control for difference in standard error of means
treatment_ohp = ohp_data%>% filter(treatment==1)
control_ohp = ohp_data %>% filter(treatment == 0 )

#Create standard error function
std <- function(x) sd(x, na.rm = TRUE)/sqrt(length(x))

#Check function and table returns the same results as if a regression was run
#lm(gender_inp ~ treatment, data = ohp_data)
 



thisworks = list(c(0.009, 0.212, 0.006, 0.016, 0.613))

#Turn into df
thisworks = as.data.frame(thisworks)

working_table = working_table%>% mutate(Standard_Error = thisworks)



kable(working_table, format = "latex", booktabs = TRUE,
        caption = "Difference in Means of Characterisitics for Treatment and Control",
        format.args = list(big.mark= ",")) %>%
    kable_styling(latex_options = c("HOLD_position","scale_down"))

```


## Question 4
>Is the balance table consistent with individuals having been randomly assigned to treatment
group and control groups? Why or why not?

The tables above show that the experiment is consistent with a random assignment for gender, race (not white), education, and cholesterol as the differences in means are not statistically significant. The difference in age is significant at the 10 percent level, which suggests that the age between the treatment and control groups may not have been properly randomized.


## Question 5 
>Estimate the compliance rate for the OHP experiment. That is, what is the effect of being
assigned to the treatment group on the probability of being enrolled in Medicaid?
Hint: For this question and question 7, you can use the same regression as in question 3,
just changing the dependent variable.

```{r pressure}
#regress ohp all ever survey on treatment
lm(data= ohp_data, ohp_all_ever_survey ~ treatment)

```
Individuals who were assigned to the treatment group were 25.36% more likely to be enrolled in Medicaid.

## Question 6
>What is the intent-to-treat (ITT) effect of the OHP experiment on health outcomes? Please
create a nicely formatted table that reports ITT estimates on 4 to 6 relevant health outcomes.
Again, part of this question is to get you to think about which 4 to 6 variables could be used
as health outcome variables.


I used diabetes diagnosis post treatment, medicine taken, hypertension post treatment, medical visits, and cholesterol. I tried to select characteristics that were important for medical use (i.e. doctors visits or prescriptions) as well as actual outcomes such as diabetes. This was done to properly evaluate the multiple impacts of the study.

```{r, results= 'asis', message=FALSE}

dia <- lm(data=ohp_data, dia_dx_post_lottery~treatment)


doc <- lm(data = ohp_data, doc_num_mod_inp~treatment)


hdp <- lm(data=ohp_data, hbp_dx_post_lottery~treatment)


meds <- lm(data=ohp_data,rx_num_mod_inp~treatment)


chl <- lm(data=ohp_data,chl_inp~treatment)



stargazer(dia, doc, hdp,meds, chl,
          title = "Overall Regression Results for Outcome Variables",
          font.size = "small",
          omit.stat = c("f", "ser"),
          column.sep.width = "1pt")


```


## Question 7
>What is the “treatment on the treated” effect (ATET) of the OHP experiment, i.e. the effect
among those who applied for Medicaid? Estimate it for every health outcome you chose
in question 6 and provide some intuition for the calculation of this estimate.

```{r, results= 'asis', message=FALSE}
#List ITT coefficients
ATET = list(c(0.009, 0.396, 0.002, 0.128, -0.642))

#List outcome
Outcome = list(c("Diabetes Post", "Medical Visits", "Hypertension Post", "Medication","Cholesterol" ))

ATET = as.data.frame(ATET)

#Create data frame
Outcome = as.data.frame(Outcome)
Outcome = Outcome %>% mutate(ATET = ATET/.2536)%>% mutate(Outcome = Outcome)
Outcome = Outcome %>% select(ATET, Outcome)
#Output table
kable(Outcome, format = "latex", booktabs = TRUE,
        caption = "ATET",
        format.args = list(big.mark= ","))



```

To calculate the ATET the coefficients in Table 4 (the intent to treat effect) were divided by the compliance rate. This was to determine the impact on treated individuals that actually participated in in the experiment.

## Question 8
>Do you have to worry about attrition bias in analyzing this data? Explain why or why not.

We should always be worried about attrition bias, as participants could theoretically move out of Oregon or stop participating. However, because the study was conducted over the relatively short period of the 10th of march 2008 to September, 30th 2009, there is not enough time for participants to develop survey fatiugue. For this reason, we do not have to be that worried about attrition bias. 





